---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

![stability-wip](https://img.shields.io/badge/stability-work_in_progress-lightgrey.svg)
[![Travis-CI Build Status](https://travis-ci.org/poissonconsulting/smbr.svg?branch=master)](https://travis-ci.org/poissonconsulting/smbr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/smbr?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/smbr)
[![codecov](https://codecov.io/gh/poissonconsulting/smbr/branch/master/graph/badge.svg)](https://codecov.io/gh/poissonconsulting/smbr)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)

# smbr

## Introduction

`smbr` (pronounced simber) is an R package to facilitate analyses using [STAN](http://mc-stan.org/about/).
It is part of the [mbr](https://github.com/poissonconsulting/mbr) family of packages.

## Installation

You can install smbr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("poissonconsulting/smbr")
```

## Demonstration

```{r, message = FALSE}
library(bauw)
library(ggplot2)
library(magrittr)
library(smbr)
```

```{r}
# define model in Stan language
model <- model("
  data {
    
    // Data lengths
      int nAnnual;
      int nObs;

    // Data
      int Annual[nObs];
      int Pairs[nObs];
      real Year[nObs];

  }
  parameters {

    // Random effects
      vector[nAnnual] bAnnual;
      real log_sAnnual;

    // Regression model
      real alpha;
      real beta1;
      real beta2;
      real beta3;

  }
  transformed parameters {
  
    real sAnnual;
    sAnnual = exp(log_sAnnual);

  }
  model {

    // Vectors of expected values
      vector[nObs] ePairs;
    
    // Priors on random effects
      log_sAnnual ~ normal(0, 10);
      bAnnual ~ normal(0, sAnnual);

    // Priors on regression coefficients
      alpha ~ normal(0, 10);
      beta1 ~ normal(0, 10);
      beta2 ~ normal(0, 10);
      beta3 ~ normal(0, 10);

    // Model
      for (i in 1:nObs) {
        ePairs[i] = exp(alpha + beta1 * Year[i] + beta2 * Year[i]^2 + 
                      beta3 * Year[i]^3 + bAnnual[Annual[i]]);
      }

      target += poisson_lpmf(Pairs | ePairs);

  }")

# add R code to modify data before running Stan
model %<>% update_model(modify_data = function(data) {
  data$nObs <- length(data$Pairs)
  data$Annual %<>% as.integer()
  data
})

# add R code to calculate derived parameters
model %<>% update_model(new_expr = "
  for (i in 1:length(Pairs)) {
    prediction[i] <- exp(alpha + beta1 * Year[i] + beta2 * Year[i]^2 + 
                       beta3 * Year[i]^3 + bAnnual[Annual[i]])
  }
  
  log_lik <- dpois(Pairs, prediction, log = TRUE)

  ")

# define data types and center year
model %<>% update_model(
  select_data = list("Pairs" = integer(), "Year*" = integer(), 
                     Annual = factor()),
  derived = "sAnnual",
  random_effects = list(bAnnual = "Annual"))

data <- bauw::peregrine
data$Annual <- factor(data$Year)

# analyse
analysis <- analyse(model, data = data)
analysis %<>% reanalyse(rhat = 1.05)

# coefficient table
coef(analysis)

# trace plots
plot(analysis)

# widely applicable information criterion (WAIC)
IC(analysis) # should be about 309.5

```
```{r, message = FALSE}
# make predictions by varying year with other predictors including the random effect of Annual held constant
year <- predict(analysis, new_data = "Year")

# plot those predictions
ggplot(data = year, aes(x = Year, y = estimate)) +
  geom_point(data = bauw::peregrine, aes(y = Pairs)) +
  geom_line() +
  geom_line(aes(y = lower), linetype = "dotted") +
  geom_line(aes(y = upper), linetype = "dotted") +
  expand_limits(y = 0)
```
## Contribution

Please report any [issues](https://github.com/poissonconsulting/smbr/issues).

[Pull requests](https://github.com/poissonconsulting/smbr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

## Inspiration

- [jaggernaut](https://github.com/poissonconsulting/jaggernaut)

## Documentation

- [STAN](http://mc-stan.org/users/documentation/)
